### 2024-06-16 code/eda-code/data_exploration_part_2.qmd
add previous scripts for historical purposes, refactoring to Python with PySpark implementation going forward

```{r}
library(readxl) #for loading Excel files
library(readr)
library(dplyr) #for data processing/cleaning
library(tidyr) #for data processing/cleaning
library(skimr) #for nice visualization of data 
library(here) #to set paths
library(tidyverse)
library(ggplot2)
library(tidycensus)
```

```{r}
getwd()
```

```{r}
# Import Data, Check Descriptive Statistics & Data Types
data_location <- here::here("starter-analysis-exercise","code","final_project","Electric_Vehicle_Population_Data.csv")
ev_data <- read.csv(data_location)
summary(ev_data)
dplyr::glimpse(ev_data)
```

```{r}
# Checking for duplicates:
duplicates <- duplicated(ev_data) # Check for duplicates
num_duplicates <- sum(duplicates) # Count of duplicates
print(num_duplicates)
```

```{r}
# Find total rows with MSRP prices
count_zero_msrp = length(which(ev_data$Base.MSRP == 0))
total_rows_count = nrow(ev_data)
total_price_data_points = total_rows_count - count_zero_msrp
print(paste0('Total rows: ', total_rows_count))
print(paste0('Total Missing MSRP Prices: ', count_zero_msrp))
print(paste0('Total Price Data Points: ', total_price_data_points))
```

```{r}
# Checking for Null values:
total_na <- sum(is.na(ev_data))
print(total_na)
```

```{r}
# Removing null values and outlier from MSRP:
ev_data_filtered <- na.omit(ev_data) # FIltered data with NA values removed
max_msrp_outlier <- ev_data_filtered %>%
  group_by(Electric.Vehicle.Type) %>%
  summarise(max = max(Base.MSRP)) 
max_msrp_outlier
filter_msrp_outlier <- ev_data_filtered %>%
  filter(Base.MSRP != 845000) %>%
  filter(Base.MSRP != 0)
filter_msrp_outlier
```

```{r}
# Create a boxplot of price distribution by electric vehicle type
p2_boxplot<-ggplot(filter_msrp_outlier, aes(x = Electric.Vehicle.Type, y = Base.MSRP)) +
  geom_boxplot() +
  labs(title = "Boxplot - Price Distribution vs. Electric Vehicle Type",
       x = "Electric Vehicle Type",
       y = "MSRP")
p2_boxplot 
```

```{r}
# Clean data for plot
price_by_state = ev_data_filtered %>% 
  group_by(Clean.Alternative.Fuel.Vehicle..CAFV..Eligibility) %>%
  filter(Base.MSRP != 0) %>%
  filter(Base.MSRP != 845000)

# Create a boxplot for price distribution by alternative fuel eligibility
p3_boxplot<-ggplot(price_by_state, aes(x = Clean.Alternative.Fuel.Vehicle..CAFV..Eligibility, y = Base.MSRP)) +
  geom_boxplot() +
  labs(title = "Boxplot - Price Distribution vs. Alt. Vehicle Eligibility",
       x = "Electric Vehicle Type",
       y = "MSRP")
p3_boxplot 
```

```{r}
# Clean data from plot
price_by_state = ev_data_filtered %>% 
  group_by(Clean.Alternative.Fuel.Vehicle..CAFV..Eligibility) %>%
  filter(Electric.Range != 0)

# Create a boxplot
p4_boxplot<-ggplot(price_by_state, aes(x = Clean.Alternative.Fuel.Vehicle..CAFV..Eligibility, y = Electric.Range)) +
  geom_boxplot() +
  labs(title = "Boxplot - Price Distribution vs. Electric Vehicle Type")
p4_boxplot 
```
```{r}
# Create a boxplot with the new categorical variable on the x-axis and height on the y-axis
p_boxplot<-ggplot(filter_msrp_outlier, aes(x = Make, y = Base.MSRP)) +
  geom_boxplot() +
  labs(title = "Boxplot - Price Distribution by Brand",
       x = "Brand",
       y = "Price")
p_boxplot
```
``` {r}
# Create a scatterplot with electric vehicle range growth over time

p_scatterplot<-ggplot(ev_data_filtered, aes(x = Model.Year, y = Electric.Range)) +
  geom_point() +
  labs(title = "Scatterplot - Electric Vehicle Range Growth over Time",
       x = "Year",
       y = "Electric Range")
plot(p_scatterplot)
# Save the scatterplot to a file
# ggsave("scatterplot.png", plot = p_scatterplot)
```

``` {r}
# Create a scatterplot - price vs. electric vehicle range correlation

p_scatterplot<-ggplot(filter_msrp_outlier, aes(x = Electric.Range, y = Base.MSRP)) +
  geom_point() +
  labs(title = "Scatterplot - Price vs. Range Correlation",
       x = "Vehicle Range",
       y = "MSRP")
plot(p_scatterplot)
```

```{r}
#For Electric.Range:
Q1 <- quantile(ev_data_filtered$Electric.Range , 0.25, na.rm = TRUE) # First Quartile
Q3 <- quantile(ev_data_filtered$Electric.Range , 0.75, na.rm = TRUE) # Third Quartile
IQR <- Q3 - Q1 # Compute Interquartile Range
lower_bound <- Q1 - 1.5 * IQR # Compute Lower Bound based on IQR
upper_bound <- Q3 + 1.5 * IQR # Compute Upper Bound based on IQR
upper_bound
```

```{r}
# Filter out outliers:
ev_data_Filtered_2 <- ev_data_filtered[ev_data_filtered$Electric.Range > lower_bound & ev_data_filtered$Electric.Range < upper_bound, ]
ev_data_Filtered_2
```