### 2024-06-22 code/eda-code/eda-ev.qmd
minor edits & begin working with geojson

---
title: "EV Exploratory Data Analysis"
date: today
output: html_document
execute: 
  error: False
  warning: False
---

# Setup

```{python}
#load needed packages. make sure they are installed.
from pandas import read_csv, set_option, merge, DataFrame, concat, read_json
import numpy as np
import matplotlib.pyplot as plt
import seaborn as sns
from sklearn.model_selection import train_test_split
from sklearn.linear_model import LinearRegression, LogisticRegression
from sklearn.preprocessing import LabelEncoder, StandardScaler
from sklearn.metrics import accuracy_score
import statsmodels.api as sm
import statsmodels
from statsmodels.graphics.gofplots import ProbPlot
from scipy import stats
import geojson
import folium
import json
set_option('display.max_columns', None)
```

Load the data.

```{python}
base_dir = '../../'
df = read_csv(base_dir + 'data/processed-data/brand_model_year_tax_data.csv')
columns_list = df.columns.tolist()
for i in columns_list:
    print(i.strip())
numeric_cols = columns_list[28:]
df.columns.tolist()
```

Vehicle Brand Aggregation
```{python}
sns.set_style('darkgrid')
zip_code_count = df.groupby(['ZIPCODE']).agg({
    'Vehicle Count': 'sum',
}).reset_index()
fig, axes = plt.subplots(2, 1, figsize=(15,15))
sns.histplot(ax=axes[0], data=zip_code_count, y=zip_code_count['Vehicle Count'])
log_zip_code_count = zip_code_count.copy()
log_zip_code_count['Vehicle Count'] = np.log(zip_code_count['Vehicle Count'])
sns.histplot(ax=axes[1], data=log_zip_code_count, y=log_zip_code_count['Vehicle Count'])
zip_code_count
```

```{python}
us_zips_df = read_csv(base_dir + 'data/raw-data/uszips.csv')
wa_zips = us_zips_df.loc[us_zips_df['state_id'] == 'WA']
wa_zips = wa_zips[['ZIPCODE', 'lat', 'lng', 'population']]
merge_zips = merge(wa_zips, zip_code_count, on='ZIPCODE')
# print(merge_zips)
test = base_dir + 'data/raw-data/counties.geojson'
with open(test) as f:
    gj = geojson.load(f)
    # print(gj['features'])
# countries = ["53"]
# new_list = []
# features = gj['features'][0]['properties']['STATEFP']
features = gj['features']
print(features[0])
# for i in features:
#     # print(i)
#     if i['properties']['STATEFP'] == '53':
#         new_list.append(i)
#     else:
#         pass
# final = json.dumps(new_list, indent=2)
# print(gj)
mapit = folium.Map(location=[122.511983, 37.77113])

for i in features:
    if i['properties']['STATEFP'] == '53':
        print(i)
        folium.Choropleth(geo_data=i['geometry']['coordinates'][0][0],
            data=merge_zips,
            columns=['lat', 'lng', 'Vehicle Count'],
            key_on='geometry.coordinates',
            fill_color='magma',
            line_color='magma'
        ).add_to(mapit)
# mapit.save('test.html')
# coordinates = []
# for zip_code in zip_code_count['ZIPCODE']:
#     print(zip_code)
#     geocode_result = geopandas.
    # geocode_result = gmaps.geocode(zip_code)
    # lat = geocode_result[0][‘geometry’][‘location’][‘lat’]
    # lng = geocode_result[0][‘geometry’][‘location’][‘lng’]
    # coordinates.append((lat, lng))
```

```{python} fig, axes = plt.subplots(2, 1, figsize=(15,15))
sns.boxplot(ax=axes[0], data=zip_code_count, x=zip_code_count['Vehicle Count'])
log_zip_code_count = zip_code_count.copy()
log_zip_code_count['Vehicle Count'] = np.log(zip_code_count['Vehicle Count'])
sns.boxplot(ax=axes[1], data=log_zip_code_count, x=log_zip_code_count['Vehicle Count'])
```

MAKE AND ELECTRIC VEHICLE TYPE
```{python}
sns.set_style('darkgrid')
make_ev_type = df.groupby(['Make']).agg({
    'Vehicle Count': 'sum',
}).reset_index()
fig, axes = plt.subplots(2, 1, figsize=(15,15))
sns.histplot(ax=axes[0], data=make_ev_type, y=make_ev_type['Vehicle Count'])
log_make_ev_type = make_ev_type.copy()
log_make_ev_type['Vehicle Count'] = np.log(make_ev_type['Vehicle Count'])
sns.histplot(ax=axes[1], data=log_make_ev_type, y=log_make_ev_type['Vehicle Count'])
```

```{python}
fig, axes = plt.subplots(2, 1, figsize=(15,15))
sns.boxplot(ax=axes[0], data=make_ev_type)
log_make_ev_type = make_ev_type.copy()
log_make_ev_type['Vehicle Count'] = np.log(make_ev_type['Vehicle Count'])
sns.boxplot(ax=axes[1], data=log_make_ev_type)
```

MAKE AND Clean Alt Fuel
```{python}
sns.set_style('darkgrid')
make_ev_type = df.groupby(['Make', 'Clean Alternative Fuel Vehicle (CAFV) Eligibility']).agg({
    'Vehicle Count': 'sum',
}).reset_index()
fig, axes = plt.subplots(2, 1, figsize=(15,15))
sns.histplot(ax=axes[0], data=make_ev_type, y=make_ev_type['Vehicle Count'])
log_make_ev_type = make_ev_type.copy()
log_make_ev_type['Vehicle Count'] = np.log(make_ev_type['Vehicle Count'])
sns.histplot(ax=axes[1], data=log_make_ev_type, y=log_make_ev_type['Vehicle Count'])
```

```{python}
fig, axes = plt.subplots(2, 1, figsize=(15,15))
sns.boxplot(ax=axes[0], data=make_ev_type, x=make_ev_type['Clean Alternative Fuel Vehicle (CAFV) Eligibility'], y=make_ev_type['Vehicle Count'])
log_make_ev_type = make_ev_type.copy()
log_make_ev_type['Vehicle Count'] = np.log(make_ev_type['Vehicle Count'])
sns.boxplot(ax=axes[1], data=log_make_ev_type, x=make_ev_type['Clean Alternative Fuel Vehicle (CAFV) Eligibility'], y=log_make_ev_type['Vehicle Count'])
```

Mean 
```{python}
brand_df = df[[
 'Residential energy tax credit amount', 'Total income amount', 'Vehicle Count'
]]

fig, axes = plt.subplots(1, 1, figsize=(15,15))
sns.heatmap(data=brand_df)
brand_df.corr()
# brand_df['Electric Range'] = np.log(brand_df['Electric Range'])
# sns.boxplot(ax=axes[1], data=brand_df, y=brand_df['Electric Range'])
```

```{python}
new_brand_df  = df.groupby(['ZIPCODE']).agg({
    'Total income amount': 'sum',
    'Electric Range': 'count'
}).reset_index(drop=True).rename(columns={
    'Electric Range': 'Vehicle Count'
})
fig, axes = plt.subplots(2, 1, figsize=(15,15))
new_brand_df['Total income amount'] = new_brand_df['Total income amount'].astype(int)
slope, intercept, r_value, p_value, std_err = stats.linregress(new_brand_df['Total income amount'],new_brand_df['Vehicle Count'])
sns.regplot(ax=axes[0], data=new_brand_df, y='Vehicle Count', x='Total income amount', 
    ci=None, 
    label=f"y={0:.1f}x+{1:.1f}\nr-sq: {r_value.round(3)}".format(slope, intercept)).legend(loc="best")
new_brand_df['Vehicle Count'] = np.log(new_brand_df['Vehicle Count'])
new_brand_df['Total income amount'] = np.log(new_brand_df['Total income amount'])
slope, intercept, r_value, p_value, std_err = stats.linregress(new_brand_df['Total income amount'],new_brand_df['Vehicle Count'])
sns.regplot(ax=axes[1],
    data=new_brand_df,
    x='Total income amount',
    y='Vehicle Count', 
    ci=None, 
    label=f"y={0:.1f}x+{1:.1f}\nr-sq: {r_value.round(3)}".format(slope, intercept)).legend(loc="best")

```

```{python}
base_dir = '../../'
df = read_csv(base_dir + 'data/processed-data/brand_model_year_tax_data.csv')
columns_list = df.columns.tolist()
for i in columns_list:
    print(i.strip())
# numeric_cols = columns_list[28:]
# df.columns.tolist()
```

ZIP CODE ONLY
```{python}
amount_brand_df = df.groupby(['ZIPCODE'])[numeric_cols].mean(numeric_only=True).reset_index()
brand_df = df.groupby(['ZIPCODE']).agg({
    'Vehicle Count': 'sum',
    'Electric Range': 'mean'
}).reset_index()
merged_df = merge(brand_df, amount_brand_df, on=['ZIPCODE'])
merged_df = merged_df.sort_values('Vehicle Count', ascending=False)

X_df = merged_df
Y_df = merged_df[['Vehicle Count']]

Y_df = np.log(Y_df)
X_df = merged_df.drop(['Vehicle Count', 'Electric Range', 'ZIPCODE'], axis=1)

X = np.array(X_df)
x = sm.add_constant(X)
Y = np.array(Y_df)
X_train, X_test, y_train, y_test = train_test_split(x,Y, 
                                   random_state=104,  
                                   test_size=0.2,
                                   shuffle=True) 
print(X_train.shape)
print(y_train.shape)
model = sm.OLS(y_train, X_train)
results = model.fit()
results.summary()

p_values = results.pvalues
p_df = DataFrame()
p_df['Feature'] = X_df.columns
p_df['p-value'] = results.pvalues[1:]
feat_list = []
p_val_list = []
for idx, row in p_df.iterrows():
    if row['p-value'] < .05:
        feat_list.append(row['Feature'])
        p_val_list.append(row['p-value'])
    else:
        continue
new_p_df = DataFrame()
new_p_df['Feature'] = feat_list
new_p_df['p-value'] = p_val_list
X = np.array(X_df[new_p_df['Feature']])
Y = np.array(Y_df)
x = sm.add_constant(X)

X_train, X_test, y_train, y_test = train_test_split(x,Y, 
                                   random_state=104,  
                                   test_size=0.2,
                                   shuffle=True) 
model2 = sm.OLS(y_train, X_train)
results2 = model2.fit()
results2.summary()

p3_values = results2.pvalues
new3 = DataFrame()
new3['Feature'] = new_p_df['Feature']
new3['p-value'] = p3_values[1:]
new3
feat_list3 = []
p_val_list3 = []
for idx, row in new3.iterrows():
    if row['p-value'] < .05:
        feat_list3.append(row['Feature'])
        p_val_list3.append(row['p-value'])
    else:
        continue
final3_df = DataFrame()
final3_df['Feature'] = feat_list3
final3_df['p-value'] = p_val_list3
final3_df
X = np.array(X_df[final3_df['Feature']])
Y = np.array(Y_df)
x = sm.add_constant(X)

X_train, X_test, y_train, y_test = train_test_split(x,Y, 
                                   random_state=104,  
                                   test_size=0.2,
                                   shuffle=True)

model3 = sm.OLS(y_train, X_train)
results3 = model3.fit()
Y_pred = results3.predict(X_train).astype(int)

features = ['constant'] + final3_df['Feature'].tolist()
X = DataFrame(X_train, columns=features)
Y = DataFrame(y_train, columns=['Vehicle Count'])
dataframe = concat([X, Y], axis=1)
dataframe
results3.summary()
features
```

```{python}
# model values
model_fitted_y = results3.fittedvalues

# model residuals
model_residuals = results3.resid

# normalized residuals
model_norm_residuals = results3.get_influence().resid_studentized_internal

# absolute squared normalized residuals
model_norm_residuals_abs_sqrt = np.sqrt(np.abs(model_norm_residuals))

# absolute residuals
model_abs_resid = np.abs(model_residuals)

# leverage, from statsmodels internals
model_leverage = results3.get_influence().hat_matrix_diag

# cook's distance, from statsmodels internals
model_cooks = results3.get_influence().cooks_distance[0]

plot_lm_1 = plt.figure()
plot_lm_1.axes[0] = sns.residplot(x=model_fitted_y, y=dataframe.columns[-1], data=dataframe,
                          lowess=True,
                          scatter_kws={'alpha': 0.5},
                          line_kws={'color': 'red', 'lw': 1, 'alpha': 0.8})

plot_lm_1.axes[0].set_title('Residuals vs Fitted')
plot_lm_1.axes[0].set_xlabel('Fitted values')
plot_lm_1.axes[0].set_ylabel('Residuals')
df = DataFrame()
df['res'] = model_residuals
df['fitted'] = model_fitted_y
df['count'] = dataframe['Vehicle Count']
# df['ZIPCODE'] = dataframe['ZIPCODE']
maxr = np.max(model_residuals)
df.sort_values('fitted', ascending=False)
```

#### Normal Q-Q Plot
This plot shows if the residuals are normally distributed. A good normal QQ plot has all of the residuals lying on or close to the red line.

```{python}
QQ = ProbPlot(model_norm_residuals)
plot_lm_2 = QQ.qqplot(line='45', alpha=0.5, color='#4C72B0', lw=1)
plot_lm_2.axes[0].set_title('Normal Q-Q')
plot_lm_2.axes[0].set_xlabel('Theoretical Quantiles')
plot_lm_2.axes[0].set_ylabel('Standardized Residuals')

# annotations
abs_norm_resid = np.flip(np.argsort(np.abs(model_norm_residuals)), 0)
abs_norm_resid_top_3 = abs_norm_resid[:3]
for r, i in enumerate(abs_norm_resid_top_3):
    plot_lm_2.axes[0].annotate(i,
                               xy=(np.flip(QQ.theoretical_quantiles, 0)[r],
                                   model_norm_residuals[i]))
```

#### Scale-Location
This plot is a way to check if the residuals suffer from non-constant variance, aka heteroscedasticity.

```{python}
plot_lm_3 = plt.figure()
plt.scatter(model_fitted_y, model_norm_residuals_abs_sqrt, alpha=0.5)
sns.regplot(
    x=model_fitted_y, 
    y=model_norm_residuals_abs_sqrt,
    scatter=False,
    ci=False,
    lowess=True,
    line_kws={'color': 'red', 'lw': 1, 'alpha': 0.8})
plot_lm_3.axes[0].set_title('Scale-Location')
plot_lm_3.axes[0].set_xlabel('Fitted values')
plot_lm_3.axes[0].set_ylabel('$\sqrt{|Standardized Residuals|}$')

# annotations
abs_sq_norm_resid = np.flip(np.argsort(model_norm_residuals_abs_sqrt), 0)
abs_sq_norm_resid_top_3 = abs_sq_norm_resid[:3]
for r, i in enumerate(abs_sq_norm_resid_top_3):
    plot_lm_3.axes[0].annotate(
        i,
        xy=(model_fitted_y[i],
        model_norm_residuals_abs_sqrt[i]))
```

#### Residuals vs. Leverage
Unlike outliers, which have an unusually large y value, leverage points have extreme x values. This may not seem so bad at face value, but it can have damaging effects on the model because the β coefficients are very sensitive to leverage points. The purpose of the Residuals vs Leverage plot is to identify these problematic observations.

```{python}
plot_lm_4 = plt.figure()
plt.scatter(model_leverage, model_norm_residuals, alpha=0.5)
sns.regplot(x=model_leverage, y=model_norm_residuals,
            scatter=False,
            ci=False,
            lowess=True,
            line_kws={'color': 'red', 'lw': 1, 'alpha': 0.8})
plot_lm_4.axes[0].set_xlim(0, np.max(model_leverage)+0.01)
plot_lm_4.axes[0].set_ylim(-3, 5)
plot_lm_4.axes[0].set_title('Residuals vs Leverage')
plot_lm_4.axes[0].set_xlabel('Leverage')
plot_lm_4.axes[0].set_ylabel('Standardized Residuals')

# annotations
leverage_top_3 = np.flip(np.argsort(model_cooks), 0)[:3]
for i in leverage_top_3:
    plot_lm_4.axes[0].annotate(i,
                                xy=(model_leverage[i],
                                    model_norm_residuals[i]))
```

```{python}
def cookdplot(model, ax=None):
    if ax is None:
        fig, ax = plt.subplots()

    cooks_d = model.get_influence().cooks_distance[0]

    ax.vlines(range(len(cooks_d)),0,cooks_d)

    # annotations
    cookd_top_3 = np.flip(np.argsort(cooks_d), 0)[:3]
    for i in cookd_top_3:
        ax.annotate(i, xy=(i, cooks_d[i]),color = 'C3')

    ax.set_title("Cook's Distance" , fontweight="bold")
    ax.set_xlabel('Obs. Number')
    ax.set_ylabel("Cook's distance")
    return ax

cookdplot(results3)
plt.show()
```

### SECOND MODEL
```{python}
# merged_df = merged_df.loc[(merged_df['Vehicle Count'] != 1.94591)]
# print(merged_df.loc[merged_df['Vehicle Count'] < 5].sort_values('Vehicle Count', ascending=False).head(30))
X_df = merged_df.copy()
Y_df = merged_df[['Vehicle Count']]
Y_df = np.log(Y_df)

X_features = [
#  'Adjust gross income (AGI) [8]',
#  'Total income amount',
 'State and local income tax refunds amount',
#  'Taxable pensions and annuities amount',
 'Number of farm returns',
 'Total statutory adjustments amount',
 'Self-employed (Keogh) retirement plans amount',
 'Total standard deduction amount',
#  'Basic standard deduction amount',
#  'Additional standard deduction amount',
#  'Total medical and dental expense deduction amount',
 'Limited state and local taxes',
 'Home mortgage from personal seller amount',
 'Alternative minimum tax amount',
 'Child and dependent care credit amount',
#  'Net premium tax credit amount',
#  'Qualified sick and family leave credit for leave taken before April 1, 2021 amount',
#  'Credited to next year’s estimated tax amount'
 ]
X_df = X_df[X_features]
feat_list = []
log_list = []
for feat in X_features:
    feat_list.append(feat)
    log_list.append(np.log(X_df[feat]))

# X_features = ['ZIPCODE',
#  'Adjust gross income (AGI) [8]',
#  'Total income amount',
#  'Total statutory adjustments amount',
#  'Home mortgage from personal seller amount',
#  'Alternative minimum tax amount']

X = X_df[X_features]

Y = Y_df
x = sm.add_constant(X)

X_train, X_test, y_train, y_test = train_test_split(x,Y, 
                                   random_state=104,  
                                   test_size=0.2,
                                   shuffle=True)

print(X_train)
model = sm.OLS(y_train, X_train)
results2mod = model.fit()
# Y_pred = results2mod.predict(X_test).astype(int)
results2mod.summary()
model1 = LinearRegression().fit(X_train, y_train)
model1.score(X_train, y_train)
# model1.score(X_test, y_test)
# results2mod.summary()

X = DataFrame(X_train, columns=['constant'] + X_features)
Y = DataFrame(y_train, columns=['Vehicle Count'])
dataframe = concat([X, Y], axis=1)
# dataframe
results2mod.summary()
# model1 = LinearRegression().fit(X_train, y_train)
# model1.score(X_train, y_train)
# model1.score(X_test, y_test)
# model.coef_
# model.intercept_
# Y_pred = model1.predict(X_test)
# plt.scatter(x=y_test, y=Y_pred)
# plt.plot(y_test, Y_pred, color='red')    
# sns.regplot(x=y_test, y=Y_pred)
```

```{python}
_# model values
model_fitted_y = results2mod.fittedvalues

# model residuals
model_residuals = results2mod.resid

# normalized residuals
model_norm_residuals = results2mod.get_influence().resid_studentized_internal

# absolute squared normalized residuals
model_norm_residuals_abs_sqrt = np.sqrt(np.abs(model_norm_residuals))

# absolute residuals
model_abs_resid = np.abs(model_residuals)

# leverage, from statsmodels internals
model_leverage = results2mod.get_influence().hat_matrix_diag
print(model_leverage)

# cook's distance, from statsmodels internals
model_cooks = results2mod.get_influence().cooks_distance[0]

plot_lm_1 = plt.figure()
plot_lm_1.axes[0] = sns.residplot(x=model_fitted_y, y=dataframe.columns[-1], data=dataframe,
                          lowess=True,
                          scatter_kws={'alpha': 0.5},
                          line_kws={'color': 'red', 'lw': 1, 'alpha': 0.8})

plot_lm_1.axes[0].set_title('Residuals vs Fitted')
plot_lm_1.axes[0].set_xlabel('Fitted values')
plot_lm_1.axes[0].set_ylabel('Residuals')
df = DataFrame()
df['res'] = model_residuals
df['fitted'] = model_fitted_y
df['count'] = dataframe['Vehicle Count']
df['index'] = df.index
maxr = np.max(model_residuals)
df.sort_values('res', ascending=True)
```

#### Normal Q-Q Plot
This plot shows if the residuals are normally distributed. A good normal QQ plot has all of the residuals lying on or close to the red line.

```{python}
QQ = ProbPlot(model_norm_residuals)
plot_lm_2 = QQ.qqplot(line='45', alpha=0.5, color='#4C72B0', lw=1)
plot_lm_2.axes[0].set_title('Normal Q-Q')
plot_lm_2.axes[0].set_xlabel('Theoretical Quantiles')
plot_lm_2.axes[0].set_ylabel('Standardized Residuals')

# annotations
abs_norm_resid = np.flip(np.argsort(np.abs(model_norm_residuals)), 0)
abs_norm_resid_top_3 = abs_norm_resid[:3]
for r, i in enumerate(abs_norm_resid_top_3):
    plot_lm_2.axes[0].annotate(i,
                               xy=(np.flip(QQ.theoretical_quantiles, 0)[r],
                                   model_norm_residuals[i]))
```

#### Scale-Location
This plot is a way to check if the residuals suffer from non-constant variance, aka heteroscedasticity.

```{python}
plot_lm_3 = plt.figure()
plt.scatter(model_fitted_y, model_norm_residuals_abs_sqrt, alpha=0.5)
sns.regplot(
    x=model_fitted_y, 
    y=model_norm_residuals_abs_sqrt,
    scatter=False,
    ci=False,
    lowess=True,
    line_kws={'color': 'red', 'lw': 1, 'alpha': 0.8})
plot_lm_3.axes[0].set_title('Scale-Location')
plot_lm_3.axes[0].set_xlabel('Fitted values')
plot_lm_3.axes[0].set_ylabel('$\sqrt{|Standardized Residuals|}$')

# annotations
abs_sq_norm_resid = np.flip(np.argsort(model_norm_residuals_abs_sqrt), 0)
abs_sq_norm_resid_top_3 = abs_sq_norm_resid[:3]
for r, i in enumerate(abs_sq_norm_resid_top_3):
    plot_lm_3.axes[0].annotate(
        i,
        xy=(model_fitted_y[i],
        model_norm_residuals_abs_sqrt[i]))
```

#### Residuals vs. Leverage
Unlike outliers, which have an unusually large y value, leverage points have extreme x values. This may not seem so bad at face value, but it can have damaging effects on the model because the β coefficients are very sensitive to leverage points. The purpose of the Residuals vs Leverage plot is to identify these problematic observations.

```{python}
plot_lm_4 = plt.figure()
plt.scatter(model_leverage, model_norm_residuals, alpha=0.5)
sns.regplot(x=model_leverage, y=model_norm_residuals,
            scatter=False,
            ci=False,
            lowess=True,
            line_kws={'color': 'red', 'lw': 1, 'alpha': 0.8})
plot_lm_4.axes[0].set_xlim(0, np.max(model_leverage)+0.01)
plot_lm_4.axes[0].set_ylim(-3, 5)
plot_lm_4.axes[0].set_title('Residuals vs Leverage')
plot_lm_4.axes[0].set_xlabel('Leverage')
plot_lm_4.axes[0].set_ylabel('Standardized Residuals')

# annotations
leverage_top_3 = np.flip(np.argsort(model_cooks), 0)[:3]
for i in leverage_top_3:
    plot_lm_4.axes[0].annotate(i,
                                xy=(model_leverage[i],
                                    model_norm_residuals[i]))
```

```{python}
def cookdplot(model, ax=None):
    if ax is None:
        fig, ax = plt.subplots()

    cooks_d = model.get_influence().cooks_distance[0]

    ax.vlines(range(len(cooks_d)),0,cooks_d)

    # annotations
    cookd_top_3 = np.flip(np.argsort(cooks_d), 0)[:3]
    for i in cookd_top_3:
        ax.annotate(i, xy=(i, cooks_d[i]),color = 'C3')

    ax.set_title("Cook's Distance" , fontweight="bold")
    ax.set_xlabel('Obs. Number')
    ax.set_ylabel("Cook's distance")
    return ax

cookdplot(results2mod)
plt.show()
```