### 2024-07-13 code/eda-code/eda-ev.qmd
comment out unfinished Python code block so QMD file renders correctly

---
title: "EV Exploratory Data Analysis"
date: today
output: html_document
execute: 
  error: False
  warning: False
---
**Electric Vehicle Data Analysis**

# Introduction:

This document is an exploratory data analysis of electric vehicle data. The data was obtained from the IRS and contains information on electric vehicles, their owners, and the tax benefits they receive. The data was cleaned and processed in R and saved as a CSV file. This document will load the data, explore it, and fit a linear regression model to predict the number of electric vehicles in a given zip code.

# Setup

```{python}
#load needed packages. make sure they are installed.
from pandas import read_csv, set_option, merge, DataFrame, concat, read_json
import numpy as np
import matplotlib.pyplot as plt
import seaborn as sns
from sklearn.model_selection import train_test_split
from sklearn.linear_model import LinearRegression, LogisticRegression
from sklearn.preprocessing import LabelEncoder, StandardScaler
from sklearn.metrics import accuracy_score
import statsmodels.api as sm
import statsmodels
from statsmodels.graphics.gofplots import ProbPlot
from scipy import stats
import geojson
import folium
import json
set_option('display.max_columns', None)
```

Load the data.

```{python}
base_dir = '../../'
df = read_csv(base_dir + 'data/processed-data/brand_model_year_tax_data.csv')
columns_list = df.columns.tolist()
for i in columns_list:
    print(i.strip())
numeric_cols = columns_list[28:]
df.columns.tolist()
```

Vehicle Brand Aggregation

```{python}
sns.set_style('darkgrid')
zip_code_count = df.groupby(['ZIPCODE']).agg({
    'Vehicle Count': 'sum',
}).reset_index()
```
The figure contains two histograms showing the distribution of vehicle counts by ZIP code. The top plot displays the original distribution, with most ZIP codes having a lower vehicle count and a few having very high counts, resulting in a skewed distribution. The bottom plot shows the log-transformed distribution, which spreads out the data and makes it easier to visualize lower counts. Both plots use a horizontal bar layout for easy comparison.

```{python}
fig, axes = plt.subplots(2, 1, figsize=(15,15))
sns.histplot(ax=axes[0], data=zip_code_count, y=zip_code_count['Vehicle Count'])
log_zip_code_count = zip_code_count.copy()
log_zip_code_count['Vehicle Count'] = np.log(zip_code_count['Vehicle Count'])
sns.histplot(ax=axes[1], data=log_zip_code_count, y=log_zip_code_count['Vehicle Count'])
zip_code_count
```

The image shows two box plots of vehicle counts by ZIP code. The top plot displays the original vehicle counts, revealing most counts below 200 with many outliers up to 1000. The bottom plot uses log-transformed vehicle counts, normalizing the data and reducing the impact of outliers. Both plots highlight the distribution, central tendency, and variability of vehicle counts across ZIP codes.

### TO DO
```{python}
# us_zips_df = read_csv(base_dir + 'data/raw-data/uszips.csv')
# wa_zips = us_zips_df.loc[us_zips_df['state_id'] == 'WA']
# wa_zips = wa_zips[['ZIPCODE', 'lat', 'lng', 'population']]
# merge_zips = merge(wa_zips, zip_code_count, on='ZIPCODE')
# # print(merge_zips)
# test = base_dir + 'data/raw-data/counties.geojson'
# with open(test) as f:
#     gj = geojson.load(f)
    # print(gj['features'])
# countries = ["53"]
# new_list = []
# features = gj['features'][0]['properties']['STATEFP']
# features = gj['features']
# print(features[0])
# for i in features:
#     # print(i)
#     if i['properties']['STATEFP'] == '53':
#         new_list.append(i)
#     else:
#         pass
# final = json.dumps(new_list, indent=2)
# print(gj)
# mapit = folium.Map(location=[122.511983, 37.77113])

# for i in features:
#     if i['properties']['STATEFP'] == '53':
#         print(i)
#         folium.Choropleth(geo_data=i['geometry']['coordinates'][0][0],
#             data=merge_zips,
#             columns=['lat', 'lng', 'Vehicle Count'],
#             key_on='geometry.coordinates',
#             fill_color='magma',
#             line_color='magma'
#         ).add_to(mapit)
# mapit.save('test.html')
# coordinates = []
# for zip_code in zip_code_count['ZIPCODE']:
#     print(zip_code)
#     geocode_result = geopandas.
    # geocode_result = gmaps.geocode(zip_code)
    # lat = geocode_result[0][‘geometry’][‘location’][‘lat’]
    # lng = geocode_result[0][‘geometry’][‘location’][‘lng’]
    # coordinates.append((lat, lng))
```

```{python} fig, axes = plt.subplots(2, 1, figsize=(15,15))
sns.boxplot(ax=axes[0], data=zip_code_count, x=zip_code_count['Vehicle Count'])
log_zip_code_count = zip_code_count.copy()
log_zip_code_count['Vehicle Count'] = np.log(zip_code_count['Vehicle Count'])
sns.boxplot(ax=axes[1], data=log_zip_code_count, x=log_zip_code_count['Vehicle Count'])
```

MAKE AND ELECTRIC VEHICLE TYPE
```{python}
sns.set_style('darkgrid')
make_ev_type = df.groupby(['Make']).agg({
    'Vehicle Count': 'sum',
}).reset_index()
fig, axes = plt.subplots(2, 1, figsize=(15,15))
sns.histplot(ax=axes[0], data=make_ev_type, y=make_ev_type['Vehicle Count'])
log_make_ev_type = make_ev_type.copy()
log_make_ev_type['Vehicle Count'] = np.log(make_ev_type['Vehicle Count'])
sns.histplot(ax=axes[1], data=log_make_ev_type, y=log_make_ev_type['Vehicle Count'])
```

```{python}
fig, axes = plt.subplots(2, 1, figsize=(15,15))
sns.boxplot(ax=axes[0], data=make_ev_type)
log_make_ev_type = make_ev_type.copy()
log_make_ev_type['Vehicle Count'] = np.log(make_ev_type['Vehicle Count'])
sns.boxplot(ax=axes[1], data=log_make_ev_type)
```

MAKE AND Clean Alt Fuel
```{python}
sns.set_style('darkgrid')
make_ev_type = df.groupby(['Make', 'Clean Alternative Fuel Vehicle (CAFV) Eligibility']).agg({
    'Vehicle Count': 'sum',
}).reset_index()
fig, axes = plt.subplots(2, 1, figsize=(15,15))
sns.histplot(ax=axes[0], data=make_ev_type, y=make_ev_type['Vehicle Count'])
log_make_ev_type = make_ev_type.copy()
log_make_ev_type['Vehicle Count'] = np.log(make_ev_type['Vehicle Count'])
sns.histplot(ax=axes[1], data=log_make_ev_type, y=log_make_ev_type['Vehicle Count'])
```

```{python}
fig, axes = plt.subplots(2, 1, figsize=(15,15))
sns.boxplot(ax=axes[0], data=make_ev_type, x=make_ev_type['Clean Alternative Fuel Vehicle (CAFV) Eligibility'], y=make_ev_type['Vehicle Count'])
log_make_ev_type = make_ev_type.copy()
log_make_ev_type['Vehicle Count'] = np.log(make_ev_type['Vehicle Count'])
sns.boxplot(ax=axes[1], data=log_make_ev_type, x=make_ev_type['Clean Alternative Fuel Vehicle (CAFV) Eligibility'], y=log_make_ev_type['Vehicle Count'])
```

Mean 
```{python}
brand_df = df[[
 'Residential energy tax credit amount', 'Total income amount', 'Vehicle Count'
]]

fig, axes = plt.subplots(1, 1, figsize=(15,15))
sns.heatmap(data=brand_df)
brand_df.corr()
# brand_df['Electric Range'] = np.log(brand_df['Electric Range'])
# sns.boxplot(ax=axes[1], data=brand_df, y=brand_df['Electric Range'])
```

```{python}
new_brand_df  = df.groupby(['ZIPCODE']).agg({
    'Total income amount': 'sum',
    'Electric Range': 'count'
}).reset_index(drop=True).rename(columns={
    'Electric Range': 'Vehicle Count'
})
fig, axes = plt.subplots(2, 1, figsize=(15,15))
new_brand_df['Total income amount'] = new_brand_df['Total income amount'].astype(int)
slope, intercept, r_value, p_value, std_err = stats.linregress(new_brand_df['Total income amount'],new_brand_df['Vehicle Count'])
sns.regplot(ax=axes[0], data=new_brand_df, y='Vehicle Count', x='Total income amount', 
    ci=None, 
    label=f"y={0:.1f}x+{1:.1f}\nr-sq: {r_value.round(3)}".format(slope, intercept)).legend(loc="best")
new_brand_df['Vehicle Count'] = np.log(new_brand_df['Vehicle Count'])
new_brand_df['Total income amount'] = np.log(new_brand_df['Total income amount'])
slope, intercept, r_value, p_value, std_err = stats.linregress(new_brand_df['Total income amount'],new_brand_df['Vehicle Count'])
sns.regplot(ax=axes[1],
    data=new_brand_df,
    x='Total income amount',
    y='Vehicle Count', 
    ci=None, 
    label=f"y={0:.1f}x+{1:.1f}\nr-sq: {r_value.round(3)}".format(slope, intercept)).legend(loc="best")

```

```{python}
base_dir = '../../'
df = read_csv(base_dir + 'data/processed-data/brand_model_year_tax_data.csv')
columns_list = df.columns.tolist()
for i in columns_list:
    print(i.strip())
# numeric_cols = columns_list[28:]
# df.columns.tolist()
```

ZIP CODE ONLY
```{python}
amount_brand_df = df.groupby(['ZIPCODE'])[numeric_cols].mean(numeric_only=True).reset_index()
brand_df = df.groupby(['ZIPCODE']).agg({
    'Vehicle Count': 'sum',
    'Electric Range': 'mean'
}).reset_index()
merged_df = merge(brand_df, amount_brand_df, on=['ZIPCODE'])
merged_df = merged_df.sort_values('Vehicle Count', ascending=False)

X_df = merged_df
Y_df = merged_df[['Vehicle Count']]

Y_df = np.log(Y_df)
X_df = merged_df.drop(['Vehicle Count', 'Electric Range', 'ZIPCODE'], axis=1)

X = np.array(X_df)
x = sm.add_constant(X)
Y = np.array(Y_df)
X_train, X_test, y_train, y_test = train_test_split(x,Y, 
                                   random_state=104,  
                                   test_size=0.2,
                                   shuffle=True) 
print(X_train.shape)
print(y_train.shape)
model = sm.OLS(y_train, X_train)
results = model.fit()
results.summary()

p_values = results.pvalues
p_df = DataFrame()
p_df['Feature'] = X_df.columns
p_df['p-value'] = results.pvalues[1:]
feat_list = []
p_val_list = []
for idx, row in p_df.iterrows():
    if row['p-value'] < .05:
        feat_list.append(row['Feature'])
        p_val_list.append(row['p-value'])
    else:
        continue
new_p_df = DataFrame()
new_p_df['Feature'] = feat_list
new_p_df['p-value'] = p_val_list
X = np.array(X_df[new_p_df['Feature']])
Y = np.array(Y_df)
x = sm.add_constant(X)

X_train, X_test, y_train, y_test = train_test_split(x,Y, 
                                   random_state=104,  
                                   test_size=0.2,
                                   shuffle=True) 
model2 = sm.OLS(y_train, X_train)
results2 = model2.fit()
results2.summary()

p3_values = results2.pvalues
new3 = DataFrame()
new3['Feature'] = new_p_df['Feature']
new3['p-value'] = p3_values[1:]
new3
feat_list3 = []
p_val_list3 = []
for idx, row in new3.iterrows():
    if row['p-value'] < .05:
        feat_list3.append(row['Feature'])
        p_val_list3.append(row['p-value'])
    else:
        continue
final3_df = DataFrame()
final3_df['Feature'] = feat_list3
final3_df['p-value'] = p_val_list3
final3_df
X = np.array(X_df[final3_df['Feature']])
Y = np.array(Y_df)
x = sm.add_constant(X)

X_train, X_test, y_train, y_test = train_test_split(x,Y, 
                                   random_state=104,  
                                   test_size=0.2,
                                   shuffle=True)

model3 = sm.OLS(y_train, X_train)
results3 = model3.fit()
Y_pred = results3.predict(X_train).astype(int)

features = ['constant'] + final3_df['Feature'].tolist()
X = DataFrame(X_train, columns=features)
Y = DataFrame(y_train, columns=['Vehicle Count'])
dataframe = concat([X, Y], axis=1)
dataframe
results3.summary()
features
```

```{python}
# model values
model_fitted_y = results3.fittedvalues

# model residuals
model_residuals = results3.resid

# normalized residuals
model_norm_residuals = results3.get_influence().resid_studentized_internal

# absolute squared normalized residuals
model_norm_residuals_abs_sqrt = np.sqrt(np.abs(model_norm_residuals))

# absolute residuals
model_abs_resid = np.abs(model_residuals)

# leverage, from statsmodels internals
model_leverage = results3.get_influence().hat_matrix_diag

# cook's distance, from statsmodels internals
model_cooks = results3.get_influence().cooks_distance[0]

plot_lm_1 = plt.figure()
plot_lm_1.axes[0] = sns.residplot(x=model_fitted_y, y=dataframe.columns[-1], data=dataframe,
                          lowess=True,
                          scatter_kws={'alpha': 0.5},
                          line_kws={'color': 'red', 'lw': 1, 'alpha': 0.8})

plot_lm_1.axes[0].set_title('Residuals vs Fitted')
plot_lm_1.axes[0].set_xlabel('Fitted values')
plot_lm_1.axes[0].set_ylabel('Residuals')
df = DataFrame()
df['res'] = model_residuals
df['fitted'] = model_fitted_y
df['count'] = dataframe['Vehicle Count']
# df['ZIPCODE'] = dataframe['ZIPCODE']
maxr = np.max(model_residuals)
df.sort_values('fitted', ascending=False)
```

#### Normal Q-Q Plot
This plot shows if the residuals are normally distributed. A good normal QQ plot has all of the residuals lying on or close to the red line.

```{python}
QQ = ProbPlot(model_norm_residuals)
plot_lm_2 = QQ.qqplot(line='45', alpha=0.5, color='#4C72B0', lw=1)
plot_lm_2.axes[0].set_title('Normal Q-Q')
plot_lm_2.axes[0].set_xlabel('Theoretical Quantiles')
plot_lm_2.axes[0].set_ylabel('Standardized Residuals')

# annotations
abs_norm_resid = np.flip(np.argsort(np.abs(model_norm_residuals)), 0)
abs_norm_resid_top_3 = abs_norm_resid[:3]
for r, i in enumerate(abs_norm_resid_top_3):
    plot_lm_2.axes[0].annotate(i,
                               xy=(np.flip(QQ.theoretical_quantiles, 0)[r],
                                   model_norm_residuals[i]))
```

#### Scale-Location
This plot is a way to check if the residuals suffer from non-constant variance, aka heteroscedasticity.

```{python}
plot_lm_3 = plt.figure()
plt.scatter(model_fitted_y, model_norm_residuals_abs_sqrt, alpha=0.5)
sns.regplot(
    x=model_fitted_y, 
    y=model_norm_residuals_abs_sqrt,
    scatter=False,
    ci=False,
    lowess=True,
    line_kws={'color': 'red', 'lw': 1, 'alpha': 0.8})
plot_lm_3.axes[0].set_title('Scale-Location')
plot_lm_3.axes[0].set_xlabel('Fitted values')
plot_lm_3.axes[0].set_ylabel('$\sqrt{|Standardized Residuals|}$')

# annotations
abs_sq_norm_resid = np.flip(np.argsort(model_norm_residuals_abs_sqrt), 0)
abs_sq_norm_resid_top_3 = abs_sq_norm_resid[:3]
for r, i in enumerate(abs_sq_norm_resid_top_3):
    plot_lm_3.axes[0].annotate(
        i,
        xy=(model_fitted_y[i],
        model_norm_residuals_abs_sqrt[i]))
```

#### Residuals vs. Leverage
Unlike outliers, which have an unusually large y value, leverage points have extreme x values. This may not seem so bad at face value, but it can have damaging effects on the model because the β coefficients are very sensitive to leverage points. The purpose of the Residuals vs Leverage plot is to identify these problematic observations.

```{python}
plot_lm_4 = plt.figure()
plt.scatter(model_leverage, model_norm_residuals, alpha=0.5)
sns.regplot(x=model_leverage, y=model_norm_residuals,
            scatter=False,
            ci=False,
            lowess=True,
            line_kws={'color': 'red', 'lw': 1, 'alpha': 0.8})
plot_lm_4.axes[0].set_xlim(0, np.max(model_leverage)+0.01)
plot_lm_4.axes[0].set_ylim(-3, 5)
plot_lm_4.axes[0].set_title('Residuals vs Leverage')
plot_lm_4.axes[0].set_xlabel('Leverage')
plot_lm_4.axes[0].set_ylabel('Standardized Residuals')

# annotations
leverage_top_3 = np.flip(np.argsort(model_cooks), 0)[:3]
for i in leverage_top_3:
    plot_lm_4.axes[0].annotate(i,
                                xy=(model_leverage[i],
                                    model_norm_residuals[i]))
```

```{python}
def cookdplot(model, ax=None):
    if ax is None:
        fig, ax = plt.subplots()

    cooks_d = model.get_influence().cooks_distance[0]

    ax.vlines(range(len(cooks_d)),0,cooks_d)

    # annotations
    cookd_top_3 = np.flip(np.argsort(cooks_d), 0)[:3]
    for i in cookd_top_3:
        ax.annotate(i, xy=(i, cooks_d[i]),color = 'C3')

    ax.set_title("Cook's Distance" , fontweight="bold")
    ax.set_xlabel('Obs. Number')
    ax.set_ylabel("Cook's distance")
    return ax

cookdplot(results3)
plt.show()
```

### SECOND MODEL
```{python}
# merged_df = merged_df.loc[(merged_df['Vehicle Count'] != 1.94591)]
# print(merged_df.loc[merged_df['Vehicle Count'] < 5].sort_values('Vehicle Count', ascending=False).head(30))
X_df = merged_df.copy()
Y_df = merged_df[['Vehicle Count']]
Y_df = np.log(Y_df)

X_features = [
#  'Adjust gross income (AGI) [8]',
#  'Total income amount',
 'State and local income tax refunds amount',
#  'Taxable pensions and annuities amount',
 'Number of farm returns',
 'Total statutory adjustments amount',
 'Self-employed (Keogh) retirement plans amount',
 'Total standard deduction amount',
#  'Basic standard deduction amount',
#  'Additional standard deduction amount',
#  'Total medical and dental expense deduction amount',
 'Limited state and local taxes',
 'Home mortgage from personal seller amount',
 'Alternative minimum tax amount',
 'Child and dependent care credit amount',
#  'Net premium tax credit amount',
#  'Qualified sick and family leave credit for leave taken before April 1, 2021 amount',
#  'Credited to next year’s estimated tax amount'
 ]
X_df = X_df[X_features]
feat_list = []
log_list = []
for feat in X_features:
    feat_list.append(feat)
    log_list.append(np.log(X_df[feat]))

# X_features = ['ZIPCODE',
#  'Adjust gross income (AGI) [8]',
#  'Total income amount',
#  'Total statutory adjustments amount',
#  'Home mortgage from personal seller amount',
#  'Alternative minimum tax amount']

X = X_df[X_features]

Y = Y_df
x = sm.add_constant(X)

X_train, X_test, y_train, y_test = train_test_split(x,Y, 
                                   random_state=104,  
                                   test_size=0.2,
                                   shuffle=True)

print(X_train)
model = sm.OLS(y_train, X_train)
results2mod = model.fit()
# Y_pred = results2mod.predict(X_test).astype(int)
results2mod.summary()
model1 = LinearRegression().fit(X_train, y_train)
model1.score(X_train, y_train)
# model1.score(X_test, y_test)
# results2mod.summary()

X = DataFrame(X_train, columns=['constant'] + X_features)
Y = DataFrame(y_train, columns=['Vehicle Count'])
dataframe = concat([X, Y], axis=1)
# dataframe
results2mod.summary()
# model1 = LinearRegression().fit(X_train, y_train)
# model1.score(X_train, y_train)
# model1.score(X_test, y_test)
# model.coef_
# model.intercept_
# Y_pred = model1.predict(X_test)
# plt.scatter(x=y_test, y=Y_pred)
# plt.plot(y_test, Y_pred, color='red')    
# sns.regplot(x=y_test, y=Y_pred)
```

```{python}
_# model values
model_fitted_y = results2mod.fittedvalues

# model residuals
model_residuals = results2mod.resid

# normalized residuals
model_norm_residuals = results2mod.get_influence().resid_studentized_internal

# absolute squared normalized residuals
model_norm_residuals_abs_sqrt = np.sqrt(np.abs(model_norm_residuals))

# absolute residuals
model_abs_resid = np.abs(model_residuals)

# leverage, from statsmodels internals
model_leverage = results2mod.get_influence().hat_matrix_diag
print(model_leverage)

# cook's distance, from statsmodels internals
model_cooks = results2mod.get_influence().cooks_distance[0]

plot_lm_1 = plt.figure()
plot_lm_1.axes[0] = sns.residplot(x=model_fitted_y, y=dataframe.columns[-1], data=dataframe,
                          lowess=True,
                          scatter_kws={'alpha': 0.5},
                          line_kws={'color': 'red', 'lw': 1, 'alpha': 0.8})

plot_lm_1.axes[0].set_title('Residuals vs Fitted')
plot_lm_1.axes[0].set_xlabel('Fitted values')
plot_lm_1.axes[0].set_ylabel('Residuals')
df = DataFrame()
df['res'] = model_residuals
df['fitted'] = model_fitted_y
df['count'] = dataframe['Vehicle Count']
df['index'] = df.index
maxr = np.max(model_residuals)
df.sort_values('res', ascending=True)
```

#### Normal Q-Q Plot
This plot shows if the residuals are normally distributed. A good normal QQ plot has all of the residuals lying on or close to the red line.

```{python}
QQ = ProbPlot(model_norm_residuals)
plot_lm_2 = QQ.qqplot(line='45', alpha=0.5, color='#4C72B0', lw=1)
plot_lm_2.axes[0].set_title('Normal Q-Q')
plot_lm_2.axes[0].set_xlabel('Theoretical Quantiles')
plot_lm_2.axes[0].set_ylabel('Standardized Residuals')

# annotations
abs_norm_resid = np.flip(np.argsort(np.abs(model_norm_residuals)), 0)
abs_norm_resid_top_3 = abs_norm_resid[:3]
for r, i in enumerate(abs_norm_resid_top_3):
    plot_lm_2.axes[0].annotate(i,
                               xy=(np.flip(QQ.theoretical_quantiles, 0)[r],
                                   model_norm_residuals[i]))
```

#### Scale-Location
This plot is a way to check if the residuals suffer from non-constant variance, aka heteroscedasticity.

```{python}
plot_lm_3 = plt.figure()
plt.scatter(model_fitted_y, model_norm_residuals_abs_sqrt, alpha=0.5)
sns.regplot(
    x=model_fitted_y, 
    y=model_norm_residuals_abs_sqrt,
    scatter=False,
    ci=False,
    lowess=True,
    line_kws={'color': 'red', 'lw': 1, 'alpha': 0.8})
plot_lm_3.axes[0].set_title('Scale-Location')
plot_lm_3.axes[0].set_xlabel('Fitted values')
plot_lm_3.axes[0].set_ylabel('$\sqrt{|Standardized Residuals|}$')

# annotations
abs_sq_norm_resid = np.flip(np.argsort(model_norm_residuals_abs_sqrt), 0)
abs_sq_norm_resid_top_3 = abs_sq_norm_resid[:3]
for r, i in enumerate(abs_sq_norm_resid_top_3):
    plot_lm_3.axes[0].annotate(
        i,
        xy=(model_fitted_y[i],
        model_norm_residuals_abs_sqrt[i]))
```

#### Residuals vs. Leverage
Unlike outliers, which have an unusually large y value, leverage points have extreme x values. This may not seem so bad at face value, but it can have damaging effects on the model because the β coefficients are very sensitive to leverage points. The purpose of the Residuals vs Leverage plot is to identify these problematic observations.

```{python}
plot_lm_4 = plt.figure()
plt.scatter(model_leverage, model_norm_residuals, alpha=0.5)
sns.regplot(x=model_leverage, y=model_norm_residuals,
            scatter=False,
            ci=False,
            lowess=True,
            line_kws={'color': 'red', 'lw': 1, 'alpha': 0.8})
plot_lm_4.axes[0].set_xlim(0, np.max(model_leverage)+0.01)
plot_lm_4.axes[0].set_ylim(-3, 5)
plot_lm_4.axes[0].set_title('Residuals vs Leverage')
plot_lm_4.axes[0].set_xlabel('Leverage')
plot_lm_4.axes[0].set_ylabel('Standardized Residuals')

# annotations
leverage_top_3 = np.flip(np.argsort(model_cooks), 0)[:3]
for i in leverage_top_3:
    plot_lm_4.axes[0].annotate(i,
                                xy=(model_leverage[i],
                                    model_norm_residuals[i]))
```

```{python}
def cookdplot(model, ax=None):
    if ax is None:
        fig, ax = plt.subplots()

    cooks_d = model.get_influence().cooks_distance[0]

    ax.vlines(range(len(cooks_d)),0,cooks_d)

    # annotations
    cookd_top_3 = np.flip(np.argsort(cooks_d), 0)[:3]
    for i in cookd_top_3:
        ax.annotate(i, xy=(i, cooks_d[i]),color = 'C3')

    ax.set_title("Cook's Distance" , fontweight="bold")
    ax.set_xlabel('Obs. Number')
    ax.set_ylabel("Cook's distance")
    return ax

cookdplot(results2mod)
plt.show()
```

### Conclusion

In this analysis, two models were compared: the initial linear regression model and a refined model with selected features. The refined model showed improved performance with higher R-squared and lower AIC values, indicating a better fit. The selection of features based on statistical significance helped in enhancing the model's predictive accuracy. Ultimately, the refined model is preferred for predicting the number of electric vehicles in a given zip code due to its better performance metrics and more robust feature selection process.