### 2024-08-06 code/analysis-code/statistical_analysis.qmd
add updated models, manuscript, and final report

---
title: "Statistical Analysis"
date: today
output: html_document
execute: 
  error: False
  warning: False
---

## Import Libraries & Data
```{python}
import pandas as pd
from sklearn.model_selection import train_test_split
import re
pd.set_option('display.max_rows', None)

# Load the dataset
data = pd.read_csv('../../data/processed-data/brand_model_year_tax_data.csv').drop(
    [
        'The State Federal Information Processing System (FIPS) code',
        'The State associated with the ZIP code',
        'Number of returns [3]',
        'City',
        'State',
        'Model Year',
        'Make',
        'Model',
        'Electric Vehicle Type',
        'Clean Alternative Fuel Vehicle (CAFV) Eligibility',
    ],
    axis = 1
)

# Remove non-letters from columns names
regex = re.compile(r"\[|\]|<", re.IGNORECASE)

data.columns = [regex.sub("", col) if any(x in str(col) for x in set(('[', ']', '<'))) else col for col in data.columns.values]

data.columns = [re.sub(r'[0-9]', '', col).strip() for col in data.columns]
data.columns = [re.sub(r'/', '_', col) for col in data.columns]

# Group by 'ZIPCODE' and sum the 'Vehicle Count'
grouped_data = data.groupby('ZIPCODE')['Vehicle Count'].sum().reset_index()

# Selecting columns L to CY (indexes 11 to 102)
features_to_use = data.columns[3:-1]
print(features_to_use)

# Ensuring correct selection of numeric columns
numeric_features_to_use = data[features_to_use].select_dtypes(include=['number']).columns

# Grouping these features by ZIPCODE and summing them
grouped_features = data.groupby('ZIPCODE')[numeric_features_to_use].sum().reset_index()

# Merging the summed Vehicle Count with the grouped features on ZIPCODE
merged_data = grouped_data.merge(grouped_features, on='ZIPCODE')

# Extracting the features (X) and target (y)
X_grouped = merged_data[numeric_features_to_use]
y_grouped = merged_data['Vehicle Count']

# Splitting the data into training and testing sets
X_train_grouped, X_test_grouped, y_train_grouped, y_test_grouped = train_test_split(
    X_grouped, y_grouped, test_size=0.2, random_state=42)

```

## Preprocess, Transform, and Build Models
```{python}
import numpy as np
import pandas as pd
from sklearn.impute import SimpleImputer
import matplotlib.pyplot as plt
import seaborn as sns
from sklearn.linear_model import LinearRegression
from sklearn.ensemble import StackingRegressor, HistGradientBoostingRegressor, RandomForestRegressor
from sklearn.tree import DecisionTreeRegressor
import xgboost as xgb
from sklearn.metrics import mean_squared_error, r2_score
import scipy.stats as stats

# Define a function to log transform all features
def log_transform_all_features(data):
    return np.log1p(data)

# Impute missing values with the median and then log transform all features
def preprocess_and_log_transform_data(X):
    imputer = SimpleImputer(strategy='median')
    X_imputed = imputer.fit_transform(X)
    X_imputed_df = pd.DataFrame(X_imputed, columns=X.columns)
    X_transformed = log_transform_all_features(X_imputed_df)
    return X_transformed

# Preprocess and log transform the training and test datasets
X_train_transformed = preprocess_and_log_transform_data(X_train_grouped)
X_test_transformed = preprocess_and_log_transform_data(X_test_grouped)

# Define the models
models = {
    'HistGradientBoosting_Regressor': HistGradientBoostingRegressor(),
    'Decision_Tree_Regressor': DecisionTreeRegressor(max_depth=5, random_state=42),
    'XGBoost_Regressor': xgb.XGBRegressor(objective='reg:squarederror', n_estimators=100, learning_rate=0.1, max_depth=5, random_state=42),
    'Random_Forest_Regressor': RandomForestRegressor(random_state=42),
}

# Function to plot Q-Q plot, residuals distribution, and predicted vs. actual
def plot_model_diagnostics(y_true, y_pred, model_name):
    residuals = y_true - y_pred

    plt.figure(figsize=(18, 10))

    # Residuals Distribution
    plt.subplot(1, 2, 1)
    sns.histplot(residuals, kde=True, bins=30)
    plt.title(f'Residuals Distribution for {model_name}')
    plt.xlabel('Residuals')
    plt.ylabel('Frequency')

    # Predicted vs Actual
    plt.subplot(1, 2, 2)
    plt.scatter(y_true, y_pred, alpha=0.5, edgecolor='k')
    plt.plot([y_true.min(), y_true.max()], [y_true.min(), y_true.max()], 'r--', lw=2)
    plt.title(f'Predicted vs. Actual for {model_name}')
    plt.xlabel('Actual Values')
    plt.ylabel('Predicted Values')
    plt.grid(True)

    plt.tight_layout()
    plt.savefig(f'../../results/output/model_results/{model_name}.png')
    plt.show()

# Train each model and plot diagnostics
for name, model in models.items():
    model.fit(X_train_transformed, y_train_grouped)
    predictions = model.predict(X_test_transformed)

    # Print model evaluation metrics
    rmse = mean_squared_error(y_test_grouped, predictions, squared=False)
    r2 = r2_score(y_test_grouped, predictions)
    print(f"{name} - RMSE: {rmse}, R-squared: {r2}")

    # Plot model diagnostics
    plot_model_diagnostics(y_test_grouped, predictions, name)
```

## Variable Importance - Top 10
```{python}
from sklearn.inspection import permutation_importance
import pandas as pd

# Train the models on the transformed training data
xgboost_model = xgb.XGBRegressor(objective='reg:squarederror', n_estimators=100, learning_rate=0.1, max_depth=5, random_state=42)
hist_gb_model = HistGradientBoostingRegressor()
dec_tree_model = DecisionTreeRegressor(max_depth=5, random_state=42)
rand_for_model = RandomForestRegressor(random_state=42)

xgboost_model.fit(X_train_transformed, y_train_grouped)
hist_gb_model.fit(X_train_transformed, y_train_grouped)
dec_tree_model.fit(X_train_transformed, y_train_grouped)
rand_for_model.fit(X_train_transformed, y_train_grouped)

# Get feature names from the transformed data
feature_names = X_train_transformed.columns

# Extract feature importances from XGBoost
dec_tree_importances = dec_tree_model.feature_importances_
dec_tree_feature_importance = pd.DataFrame({
    'Feature': feature_names,
    'Importance': dec_tree_importances
}).sort_values(by='Importance', ascending=False)

# Extract feature importances from XGBoost
xgboost_importances = xgboost_model.feature_importances_
xgboost_feature_importance = pd.DataFrame({
    'Feature': feature_names,
    'Importance': xgboost_importances
}).sort_values(by='Importance', ascending=False)

# Compute feature importances using permutation importance for HistGradientBoostingRegressor
hist_gb_importances = permutation_importance(hist_gb_model, X_test_transformed, y_test_grouped, n_repeats=10, random_state=42)
hist_gb_feature_importance = pd.DataFrame({
    'Feature': feature_names,
    'Importance': hist_gb_importances.importances_mean
}).sort_values(by='Importance', ascending=False)

# Extract feature importances from XGBoost
rand_for_importances = rand_for_model.feature_importances_
rand_for_feature_importance = pd.DataFrame({
    'Feature': feature_names,
    'Importance': rand_for_importances
}).sort_values(by='Importance', ascending=False)

# Display the top 10 important features from each model
print("Top 10 Important Features from XGBoost:")
print(xgboost_feature_importance.head(10))

print("\nTop 10 Important Features from HistGradientBoostingRegressor (Permutation Importance):")
print(hist_gb_feature_importance.head(10))

print("\nTop 10 Important Features from Decision Tree:")
print(dec_tree_feature_importance.head(10))

print("\nTop 10 Important Features from Random Forest:")
print(rand_for_feature_importance.head(10))

rand_for_feature_importance_list = rand_for_feature_importance['Feature'].head(10).tolist()
print(rand_for_feature_importance_list)

# Plotting function
def plot_feature_importance(importances, model_name):
    plt.figure(figsize=(15, 10))
    plt.barh(importances['Feature'], importances['Importance'], color='skyblue')
    plt.xlabel('Importance')
    plt.ylabel('Feature')
    plt.title(f'Feature Importance - {model_name}')
    plt.gca().invert_yaxis()  # Invert y-axis to have the most important feature at the top
    plt.grid(True)
    plt.tight_layout()
    plt.savefig(f'../../results/output/variable_importance/{model_name}.png')

# Plotting feature importance for XGBoost
plot_feature_importance(xgboost_feature_importance.head(10), 'XGBoost')

# Plotting feature importance for HistGradientBoostingRegressor (Permutation Importance)
plot_feature_importance(hist_gb_feature_importance.head(10), 'HistGradientBoostingRegressor (Permutation Importance)')

# Plotting feature importance for HistGradientBoostingRegressor (Permutation Importance)
plot_feature_importance(rand_for_feature_importance.head(10), 'Random Forest Regressor')

# Plotting feature importance for HistGradientBoostingRegressor (Permutation Importance)
plot_feature_importance(dec_tree_feature_importance.head(10), 'Decision Tree Regressor')
```

## Compare Models
```{python}

from sklearn.metrics import mean_squared_error, mean_absolute_error, r2_score
import pandas as pd

# Function to evaluate models and compare various metrics
def compare_model_metrics(models, X_train, y_train, X_test, y_test):
    results = []

    for name, model in models.items():
        # Train the model
        model.fit(X_train, y_train)
        # Predict on the test set
        predictions = model.predict(X_test)
        # Calculate metrics
        rmse = mean_squared_error(y_test, predictions, squared=False)
        r2 = r2_score(y_test, predictions)
        mae = mean_absolute_error(y_test, predictions)
        mse = mean_squared_error(y_test, predictions)
        # Store the results
        results.append({
            'Model': name,
            'RMSE': rmse,
            'R-squared': r2,
            'MAE': mae,
            'MSE': mse
        })

    # Convert the results to a DataFrame for easy comparison
    return pd.DataFrame(results)

# Models to Compare
models_to_compare = {
    'XGBoost_Regressor': xgboost_model,  # Assuming xgboost_model is already defined
    'Decision_Tree_Regressor': DecisionTreeRegressor(max_depth=5, random_state=42),
    'HistGradientBoosting_Regressor': hist_gb_model,  # Assuming hist_gb_model is already defined
    'Random_Forest_Regressor': RandomForestRegressor(n_estimators=100, random_state=42),
}

# Evaluate and compare the models
model_comparison_metrics_df = compare_model_metrics(models_to_compare, X_train_transformed, y_train_grouped, X_test_transformed, y_test_grouped)

# Display the comparison results
print(model_comparison_metrics_df)
model_comparison_metrics_df.to_csv('../../results/output/model_comparison.csv', index=False)

```

## Visualize Model Comparison
```{python}
import matplotlib.pyplot as plt

# Function to visualize the model performance differences
def plot_model_performance_comparison(df, metric):
    plt.figure(figsize=(10, 6))
    df_sorted = df.sort_values(by=metric, ascending=False)
    plt.barh(df_sorted['Model'], df_sorted[metric], color='skyblue')
    plt.xlabel(metric)
    plt.ylabel('Model')
    plt.title(f'Model Comparison by {metric}')
    plt.gca().invert_yaxis()  # Invert y-axis to have the best performing model on top
    plt.grid(True)
    plt.savefig(f'../../results/output/model_comparison/{metric}.png')
    plt.show()

# Visualize RMSE comparison
plot_model_performance_comparison(model_comparison_metrics_df, 'RMSE')

# Visualize R-squared comparison
plot_model_performance_comparison(model_comparison_metrics_df, 'R-squared')

# Visualize MAE comparison
plot_model_performance_comparison(model_comparison_metrics_df, 'MAE')

# Visualize MSE comparison
plot_model_performance_comparison(model_comparison_metrics_df, 'MSE')
```

## Create Coefficients DataFrame
```{python}
# Import necessary libraries
import pandas as pd
import numpy as np

# Function to extract model coefficients and intercepts
def get_model_coefficients_and_intercepts(model, model_name, feature_names):
    if hasattr(model, 'coef_'):
        # Linear models and similar that have a coef_ attribute
        coefficients = model.coef_
    elif hasattr(model, 'feature_importances_'):
        # Tree-based models with feature_importances_ attribute
        coefficients = model.feature_importances_
    else:
        coefficients = np.nan  # Placeholder for models without accessible coefficients

    coefficients_df = pd.DataFrame({
        'Feature': feature_names,
        'Coefficient': coefficients,
        'Model': model_name
    })

    return pd.concat([coefficients_df], ignore_index=True)

# Get feature names
feature_names = X_train_transformed.columns

# Initialize models (assuming they are already trained)
models = {
    'XGBoost_Regressor': xgboost_model,
    'Decision_Tree_Regressor': dec_tree_model,
    'HistGradientBoosting_Regressor': hist_gb_model,
    'Random_Forest_Regressor': rand_for_model,
}

# Collect coefficients for each model
coefficients_with_intercepts_list = []

for model_name, model in models.items():
    coefficients_with_intercepts_list.append(get_model_coefficients_and_intercepts(model, model_name, feature_names))

# Combine all coefficients into one DataFrame
coefficients_df = pd.concat(coefficients_with_intercepts_list, ignore_index=True)

# Display the coefficients DataFrame
coefficients_df = coefficients_df.dropna()
coefficients_df.to_csv('../../results/output/model_coefficients.csv', index=False)
```