### 2024-08-06 code/eda-code/eda-ev-tax.qmd

---
title: "EV Exploratory Data Analysis"
date: today
output: html_document
execute: 
  error: False
  warning: False
---
**Electric Vehicle Data Analysis**

# Introduction:

This document is an exploratory data analysis of electric vehicle data. The data was obtained from the IRS and contains information on electric vehicles, their owners, and the tax benefits they receive. The data was cleaned and processed in R and saved as a CSV file. This document will load the data, explore it, and fit a linear regression model to predict the number of electric vehicles in a given zip code.

# Setup

```{python}
import pandas as pd
from sklearn.model_selection import train_test_split
import re

# Load the dataset
data = pd.read_csv('../../data/processed-data/brand_model_year_tax_data.csv').drop(
    [
        'The State Federal Information Processing System (FIPS) code',
        'The State associated with the ZIP code',
        'Number of returns [3]',
        'City',
        'State',
        'Model Year',
        'Make',
        'Model',
        'Electric Vehicle Type',
        'Clean Alternative Fuel Vehicle (CAFV) Eligibility',
    ],
    axis = 1
)

# Remove non-letters from columns names
regex = re.compile(r"\[|\]|<", re.IGNORECASE)

data.columns = [regex.sub("", col) if any(x in str(col) for x in set(('[', ']', '<'))) else col for col in data.columns.values]

data.columns = [re.sub(r'[0-9]', '', col).strip() for col in data.columns]
data.columns = [re.sub(r'/', '_', col) for col in data.columns]

# Group by 'ZIPCODE' and sum the 'Vehicle Count'
grouped_data = data.groupby('ZIPCODE')['Vehicle Count'].sum().reset_index()

# Selecting columns L to CY (indexes 11 to 102)
features_to_use = data.columns[2:]

# Ensuring correct selection of numeric columns
numeric_features_to_use = data[features_to_use].select_dtypes(include=['number']).columns

# Grouping these features by ZIPCODE and summing them
grouped_features = data.groupby('ZIPCODE')[numeric_features_to_use].sum().reset_index().drop('Vehicle Count', axis = 1)

# Merging the summed Vehicle Count with the grouped features on ZIPCODE
merged_data = grouped_data.merge(grouped_features, on='ZIPCODE')

# Extracting the features (X) and target (y)
X_grouped = merged_data[numeric_features_to_use]
y_grouped = merged_data['Vehicle Count']

# Splitting the data into training and testing sets
X_train_grouped, X_test_grouped, y_train_grouped, y_test_grouped = train_test_split(
    X_grouped, y_grouped, test_size=0.3, random_state=42)

```

```{python}
import numpy as np
import pandas as pd
from sklearn.impute import SimpleImputer
import matplotlib.pyplot as plt
import seaborn as sns

# Define a function to log transform all features
def log_transform_all_features(data):
    return np.log(data)

# Impute missing values with the median and then log transform all features
def preprocess_and_log_transform_data(X):
    imputer = SimpleImputer(strategy='median')
    X_imputed = imputer.fit_transform(X)
    X_imputed_df = pd.DataFrame(X_imputed, columns=X.columns)
    X_transformed = log_transform_all_features(X_imputed_df)
    return X_transformed

# Preprocess and log transform the training and test datasets
X_train_transformed = preprocess_and_log_transform_data(X_train_grouped)
X_test_transformed = preprocess_and_log_transform_data(X_test_grouped)

# Plot histogram for each feature in the transformed dataset
def plot_transformed_feature_histograms(data, data1):

    for feature in data.columns:
        fig, ax = plt.subplots(1, 2, figsize=(10,10))
        ax1 = plt.subplot(2,1,1)
        ax1 = sns.histplot(data[feature], kde=True, bins=50)
        ax2 = plt.subplot(2,1,2)
        ax2 = sns.histplot(data1[feature], kde=True, bins=50)
        ax1.title.set_text(f'Histogram of {feature}')
        ax2.title.set_text(f'Log-Transformed Histogram of {feature}')
        plt.xlabel(feature)
        plt.ylabel('Frequency')
        plt.grid(True)
        plt.savefig(f'../../results/figures/histograms/{feature}_histogram.png')

# Plot the histograms for the log-transformed features
plot_transformed_feature_histograms(X_train_grouped, X_train_transformed)
```

```{python}
# Plot boxplot for each feature in the transformed dataset
def plot_transformed_feature_boxplots(data, data1):

    for feature in data.columns:
        fig, ax = plt.subplots(1, 2, figsize=(10,10))
        ax1 = plt.subplot(2,1,1)
        ax1 = sns.boxplot(data[feature])
        ax2 = plt.subplot(2,1,2)
        ax2 = sns.boxplot(data1[feature])
        ax1.title.set_text(f'Boxplot of {feature}')
        ax2.title.set_text(f'Log-Transformed Boxplot of {feature}')
        plt.xlabel(feature)
        plt.ylabel('Frequency')
        plt.grid(True)
        plt.tight_layout()
        plt.savefig(f'../../results/figures/boxplots/{feature}_boxplot.png')

# Plot the boxplots for the log-transformed features
plot_transformed_feature_boxplots(X_train_grouped, X_train_transformed)
```