### 2024-08-06 aboutme.qmd
add first site iteration

---
title: "About Me"
---

![The Real Family](IMG_3544.jpg){height=400 fig-alt="The Real Family" fig-align="center"}

#### **Personal Background**
My name is Collin Real. I'm 27 years old, and I grew up in Saint Hedwig, Texas. My hobbies include fitness, technology, and hanging out with my friends and family. I also breed Persian cats.

#### **Education/Career**
I have a Bachelor's and Master's Degree in Accounting, and worked as a Tax Accountant for 3-4 years. As I was finishing my accounting degrees, my friend introduced me to programming, and I instantly knew I was headed for a career change. I researched how Python could be used in accounting/finance, discovering the power of automation, building websites, and data science. To gain experience, I used YouTube, GitHub, and many other websites to learn from projects created by others. After adding this experience to my resume, I joined the Macrospace Analytics team at H-E-B in 2023 as an analyst, performing typical analyst as well as data science work. I hope to transition fully to data science work once I complete the MSDA program.

#### **Storytelling w/ Data**
{{< video https://www.youtube.com/watch?v=8EMW7io4rSI >}}