### 2024-08-06 code/eda-code/eda-ev-dataset.qmd

```{r}
library(readxl) #for loading Excel files
library(readr)
library(dplyr) #for data processing/cleaning
library(tidyr) #for data processing/cleaning
library(skimr) #for nice visualization of data 
library(here) #to set paths
library(tidyverse)
library(ggplot2)
library(tidycensus)
```

# Introduction
This document is an exploratory data analysis of electric vehicle data. The data was obtained from the IRS and contains information on electric vehicles, their owners, and the tax benefits they receive. The data was cleaned and processed in R and saved as a CSV file. This document will load the data, explore it, and fit a linear regression model to predict the number of electric vehicles in a given zip code.

```{r}
# Import Data, Check Descriptive Statistics & Data Types
data_location <- here::here("data","raw-data","Electric_Vehicle_Population_Data.csv")
ev_data <- read.csv(data_location)
summary(ev_data)
dplyr::glimpse(ev_data)
```

```{r}
# Checking for duplicates:
duplicates <- duplicated(ev_data) # Check for duplicates
num_duplicates <- sum(duplicates) # Count of duplicates
print(num_duplicates)
```

```{r}
# Find total rows with MSRP prices
count_zero_msrp = length(which(ev_data$Base.MSRP == 0))
total_rows_count = nrow(ev_data)
total_price_data_points = total_rows_count - count_zero_msrp
print(paste0('Total rows: ', total_rows_count))
print(paste0('Total Missing MSRP Prices: ', count_zero_msrp))
print(paste0('Total Price Data Points: ', total_price_data_points))
```

```{r}
# Checking for Null values:
total_na <- sum(is.na(ev_data))
print(total_na)
```

```{r}
# Removing null values and outlier from MSRP:
ev_data_filtered <- na.omit(ev_data) # FIltered data with NA values removed
max_msrp_outlier <- ev_data_filtered %>%
  group_by(Electric.Vehicle.Type) %>%
  summarise(max = max(Base.MSRP)) 
max_msrp_outlier
filter_msrp_outlier <- ev_data_filtered %>%
  filter(Base.MSRP != 845000) %>%
  filter(Base.MSRP != 0)
```

```{r}
# Create a boxplot of price distribution by electric vehicle type
p2_boxplot<-ggplot(filter_msrp_outlier, aes(x = Electric.Vehicle.Type, y = Base.MSRP)) +
  geom_boxplot() +
  labs(title = "Boxplot - Price Distribution vs. Electric Vehicle Type",
       x = "Electric Vehicle Type",
       y = "MSRP")
p2_boxplot 
```

```{r}
# Clean data for plot
price_by_state = ev_data_filtered %>% 
  group_by(Clean.Alternative.Fuel.Vehicle..CAFV..Eligibility) %>%
  filter(Base.MSRP != 0) %>%
  filter(Base.MSRP != 845000)

# Create a boxplot for price distribution by alternative fuel eligibility
p3_boxplot<-ggplot(price_by_state, aes(x = Clean.Alternative.Fuel.Vehicle..CAFV..Eligibility, y = Base.MSRP)) +
  geom_boxplot() +
  labs(title = "Boxplot - Price Distribution vs. Alt. Vehicle Eligibility",
       x = "Electric Vehicle Type",
       y = "MSRP")
p3_boxplot 
```

```{r}
# Clean data from plot
price_by_state = ev_data_filtered %>% 
  group_by(Clean.Alternative.Fuel.Vehicle..CAFV..Eligibility) %>%
  filter(Electric.Range != 0)

# Create a boxplot
p4_boxplot<-ggplot(price_by_state, aes(x = Clean.Alternative.Fuel.Vehicle..CAFV..Eligibility, y = Electric.Range)) +
  geom_boxplot() +
  labs(title = "Boxplot - Price Distribution vs. Electric Vehicle Type")
p4_boxplot 
```
```{r}
# Create a boxplot with the new categorical variable on the x-axis and height on the y-axis
p_boxplot<-ggplot(filter_msrp_outlier, aes(x = Make, y = Base.MSRP)) +
  geom_boxplot() +
  labs(title = "Boxplot - Price Distribution by Brand",
       x = "Brand",
       y = "Price")
p_boxplot
```
``` {r}
# Create a scatterplot with electric vehicle range growth over time

p_scatterplot<-ggplot(ev_data_filtered, aes(x = Model.Year, y = Electric.Range)) +
  geom_point() +
  labs(title = "Scatterplot - Electric Vehicle Range Growth over Time",
       x = "Year",
       y = "Electric Range")
plot(p_scatterplot)
# Save the scatterplot to a file
# ggsave("scatterplot.png", plot = p_scatterplot)
```