### 2024-08-06 code/eda-code/eda-ev-tax.qmd

---
title: "Log-Transform Predictors"
date: today
output: html_document
execute: 
  error: False
  warning: False
---
**Electric Vehicle Data Analysis**

# Introduction
Apply log transformation to our feature

# Setup

```{python}
import pandas as pd
from sklearn.model_selection import train_test_split
import re

# Load the dataset
data = pd.read_csv('../../data/processed-data/brand_model_year_tax_data.csv').drop(
    [
        'The State Federal Information Processing System (FIPS) code',
        'The State associated with the ZIP code',
        'Number of returns [3]',
        'City',
        'State',
        'Model Year',
        'Make',
        'Model',
        'Electric Vehicle Type',
        'Clean Alternative Fuel Vehicle (CAFV) Eligibility',
    ],
    axis = 1
)

# Remove non-letters from columns names
regex = re.compile(r"\[|\]|<", re.IGNORECASE)

data.columns = [regex.sub("", col) if any(x in str(col) for x in set(('[', ']', '<'))) else col for col in data.columns.values]

data.columns = [re.sub(r'[0-9]', '', col).strip() for col in data.columns]
data.columns = [re.sub(r'/', '_', col) for col in data.columns]

# Group by 'ZIPCODE' and sum the 'Vehicle Count'
grouped_data = data.groupby('ZIPCODE')['Vehicle Count'].sum().reset_index()

# Selecting columns L to CY (indexes 11 to 102)
features_to_use = data.columns[2:]

# Ensuring correct selection of numeric columns
numeric_features_to_use = data[features_to_use].select_dtypes(include=['number']).columns

# Grouping these features by ZIPCODE and summing them
grouped_features = data.groupby('ZIPCODE')[numeric_features_to_use].sum().reset_index().drop('Vehicle Count', axis = 1)

# Merging the summed Vehicle Count with the grouped features on ZIPCODE
merged_data = grouped_data.merge(grouped_features, on='ZIPCODE')

# Extracting the features (X) and target (y)
X_grouped = merged_data[numeric_features_to_use]
y_grouped = merged_data['Vehicle Count']

# Splitting the data into training and testing sets
X_train_grouped, X_test_grouped, y_train_grouped, y_test_grouped = train_test_split(
    X_grouped, y_grouped, test_size=0.2, random_state=42)

merged_data.to_csv('test.csv', index=False)
```
<!-- 
```{python}
import matplotlib.pyplot as plt

# Scatter plots to understand relationship between predictors and independent variable
for i, feature in enumerate(features_to_use):
    plt.figure(figsize=(10,5))
    plt.scatter(X_grouped[feature], y_grouped, c="lightblue", edgecolors="blue")
    plt.title(f'{feature} vs. Vehicle Count Relationship')
    plt.xlabel(feature)
    plt.ylabel('Vehicle Count')
    # plt.grid(True)
    plt.savefig(f'../../results/figures/scatterplots/{feature}_scatterplot.png')

```

```{python}
import numpy as np
import pandas as pd
from sklearn.impute import SimpleImputer
import matplotlib.pyplot as plt
import seaborn as sns

# Define a function to log transform all features
def log_transform_all_features(data):
    return np.log(data)

# Impute missing values with the median and then log transform all features
def preprocess_and_log_transform_data(X):
    imputer = SimpleImputer(strategy='median')
    X_imputed = imputer.fit_transform(X)
    X_imputed_df = pd.DataFrame(X_imputed, columns=X.columns)
    X_transformed = log_transform_all_features(X_imputed_df)
    return X_transformed

# Preprocess and log transform the training and test datasets
X_transformed = preprocess_and_log_transform_data(X_grouped)

# Plot histogram for each feature in the transformed dataset
def plot_transformed_feature_histograms(data, data1):

    for feature in data.columns:
        fig, ax = plt.subplots(1, 2, figsize=(10,5))
        ax1 = plt.subplot(2,1,1)
        ax1 = sns.histplot(data[feature], kde=True, bins=50)
        ax2 = plt.subplot(2,1,2)
        ax2 = sns.histplot(data1[feature], kde=True, bins=50)
        ax1.title.set_text(f'Original vs. Log-Transformed: {feature}')
        # ax2.title.set_text(f'Log-Transformed Histogram of {feature}')
        plt.xlabel(feature)
        plt.ylabel('Frequency')
        plt.grid(True)
        plt.savefig(f'../../results/figures/histograms/{feature}_histogram.png')

# Plot the histograms for the log-transformed features
plot_transformed_feature_histograms(X_grouped, X_transformed)
```

```{python}
# Plot boxplot for each feature in the transformed dataset
def plot_transformed_feature_boxplots(data, data1):

    for feature in data.columns:
        fig, ax = plt.subplots(1, 2, figsize=(10,5))
        ax1 = plt.subplot(2,1,1)
        ax1 = sns.boxplot(data[feature])
        ax2 = plt.subplot(2,1,2)
        ax2 = sns.boxplot(data1[feature])
        ax1.title.set_text(f'Original vs. Log-Transformed Boxplot of {feature}')
        # ax2.title.set_text(f'Log-Transformed Boxplot of {feature}')
        plt.xlabel(feature)
        plt.ylabel('Frequency')
        plt.grid(True)
        plt.tight_layout()
        plt.savefig(f'../../results/figures/boxplots/{feature}_boxplot.png')

# Plot the boxplots for the log-transformed features
plot_transformed_feature_boxplots(X_grouped, X_transformed)
``` -->